Supplement: Analysis and code for 'Testing Perceptual Accuracy in Charts using Surveys

### 2023-07-14 analysis/analysis.qmd
number of participants and responses

Both of these factors have been shown [@hughesJustNoticeableDifferences2001; @luModelingJustNoticeable2022] to be beneficial for an accurate assessment of small differences between bars.

@@

We use the results from rounds 1 and 2 to assess perceptual accuracy in the vertical stacked barchart (as shown in @fig-tasks-12). 
@fig-designs shows the two designs shown to participants of Round 3. The survey was administered using a split design on the type of visualization, but all participants were asked to compare the size of two unaligned tiles (A and B) and two aligned tiles (C and D). Roughly half ($n_1 = $ `r n_tall` for an efficient sample size of `r n_tall_eff`) of the round 6 panelists were shown the horizontal tall barchart on the left, the other participants ($n_2 = $ `r n_wide`; efficient $n_2^{(eff)} = $ `r n_wide_eff`) saw the horizontal wide barchart.

@@

>>>>

Both of these factors have been shown [@luModelingJustNoticeable2022] to be beneficial for an accurate assessment of small differences between bars.

@@

We use the results from rounds 1 and 2 to assess perceptual accuracy in the vertical stacked barchart (as shown in @fig-tasks-12). 
@fig-designs shows the two designs shown to participants of Round 3. The survey was administered using a split design on the type of visualization, but all participants were asked to compare the size of two unaligned tiles (A and B) and two aligned tiles (C and D). Roughly half ($n_1 = $ `r n_tall` for an efficient sample size of `r round(n_tall_eff,1)`) of the round 6 panelists were shown the horizontal tall barchart on the left, the other participants ($n_2 =$   `r n_wide`; efficient $n_2^{(eff)} =$    `r round(n_wide_eff,1)`) saw the horizontal wide barchart.

@@

```{r combine-123}
lambda_1_2 <- calculate_lambda(round1_adj$weight, round2_adj$weight)

round_12 <- round1_adj %>% 
            mutate(weight = lambda_1_2*weight/sum(weight),
                   round = "Round 1") %>% 
            dplyr::select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD,
                   ab_certain, cd_certain, ab_zoom, cd_zoom, round, weight) %>%
    rbind(round2_adj %>% 
            mutate(weight = (1-lambda_1_2)*weight/sum(weight),
                   round = "Round 2") %>% 
            dplyr::select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD, 
                   ab_certain, cd_certain, ab_zoom, cd_zoom, round, weight))

lambda_12_3 <- calculate_lambda(round_12$weight, round3_adj$weight)

round_123 <- round_12 %>% 
              mutate(weight = lambda_12_3*weight/sum(weight),
                     task = "Vertical") %>% 
  rbind(
    round3_adj %>% 
      mutate(
        weight = (1-lambda_12_3)*weight/sum(weight),
        round = "Round 3",
        CorrectAB = ab == "B is bigger",
        CorrectCD = cd == "D is bigger",
        CorrectABCD = paste(CorrectAB, CorrectCD),
        task = ifelse(P_EXP==1, "Horizontal", "Horizontal wide")
       ) %>% 
        dplyr::select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD,
                   ab_certain, cd_certain, ab_zoom, cd_zoom, round, task, weight) 
  )

round_123$weight <- neff(round_123$weight)/sum(round_123$weight)*round_123$weight 
# does not affect the further analysis, but helps with the interpretation of results

round_123_long <- round_123 %>% 
  pivot_longer(CorrectAB:CorrectCD, names_to = "Question", values_to = "Response")

survey_123 <- 
  survey::svydesign(
  data = round_123_long, 
  ids = ~CaseId, weights = ~weight)
```

We combine rounds 1, 2 and 3 as shown in @tbl-combine-123 to incorporate all relevant responses to evaluate the three different designs of charts in terms of their accuracy in assessing small differences in tile sizes.

```{r tbl-combine-123}
#| label: tbl-combine-123
#| tbl-cap: Table of the number of effective responses by type of visualization and round

lambdas <- round_123_long %>% group_by(round, task) %>% 
  summarize(
    lambda = calculate_lambda(
      round_123_long$weight[round_123_long$task==task[1] & round_123_long$round==round[1]],
      round_123_long$weight[!(round_123_long$task==task[1] & round_123_long$round==round[1])])
  )

dframe <- round_123_long %>% group_by(round, task) %>% 
  summarize(
    nominal = n(),
    effective = round(neff(weight),1),
    participants = length(unique(CaseId))
  ) %>% left_join(
    lambdas, by = c("round", "task")
  )

dframe %>% 
  group_by(round) %>%
  gt() %>%
  tab_header("Number of (effective) responses by task") %>%
  cols_label(
    nominal = md("number of responses:<br>nominal"),
    effective = md("number of responses:<br>effective"),
    participants = md("number of participants"),
    lambda = md("λ"),
  ) %>%
  fmt_number(
    everything(), 
    decimals = 0,
    use_seps=TRUE
  ) %>%
  fmt_number(
    columns = effective,
    decimals = 1
  ) %>%
  fmt_number(
    columns = c(lambda),
    decimals = 4
  ) %>%
  tab_options(
    column_labels.font.weight = "bold",
    row_group.font.weight = "bold"
  )
```
